Dirac neutrino mass generation from Majorana messenger

### 2019-08-02 draft.tex
beta version of introduction

%Features of the new propossed solution
In general, solutions for light Dirac neutrino masses require a
continuous symmetry to guarantee their Diracness. This symmetry is
usually identified as the $\operatorname{U}(1)_{B-L}$.  Additionally,
ad hoc discrete symmetries are invoked to forbid tree level Dirac or
Majorana mass terms for the light right-handed
neutrinos~\cite{Roncadelli:1983ty,Han:2018zcn,Wang:2017mcy}. 
%tree-level without adhoc symmetries
However, tree-level Dirac type-I seesaw with proper choices for the
$\operatorname{U}(1)_{B-L}$ charges have been shown to be consistent
without require any extra ad-hoc discrete symmetry~\cite{Ma:2014qra}.
%one-loop 
In recent works, it has been shown that even for one-loop Dirac
neutrino masses, it is possible to have $\operatorname{U}(1)_{B-L}$ as
the only extra symmetry beyond the SM~\cite{Calle:2018ovc,
  Bonilla:2018ynb, Saad:2019bqf}\footnote{
%extra-loops
  For extensions with only extra scalars, solutions have been found with two and three loops~\cite{Saad:2019bqf}}.
%conexion with dark matter
As a bonus in this case, the new scalars and fermions circulating the
loop can be dark matter candidates with the stability of the lightest
of them guaranteed by the very same continuous symmetry.
% remark
We focus here in solutions for the radiative Dirac type-I seesaw with
Majorana mediators, which have only an extra symmetry responsible for the
Diracness of the light neutrinos, the absence of any tree-level
contribution and the existence of a dark sector constituted by the
particles circulating the loop.

%conexion with DM
In fact, another evidence that the SM is not a complete theory is the
missing matter content of the universe, which is known as dark matter
(DM).
% DM as particle
The main proposals that explain DM as a particle are given in
Ref.~\cite{Bertone:2004pz}.
% too many possibilities
Without evidence of DM as a particle there is not a clear path to pin
out the DM properties nor the possible heavier companions of some
extended sector. 
% connect with other phenomena
Linking the dark sector to other specific phenomenology allows to
reduce the arbitrariness in the model building.
% loop
In our construction, the dark sector is related to the heavy sector
responsible of the lightness of the neutrinos and the same symmetry
which guarantees the lightness of the Dirac neutrinos is the
responsible of the stability of the lightest dark particle (LDP).
% new particle prediction
The number of specific models is restricted.

@@

%
\begin{table}
  \centering
  \begin{tabular}{|l|l|c|c|}
    \hline  
    Particles     & $\left( SU(2)_L, U(1)_Y \right)$ & $U(1)_{L}$& $Z_2$\\ \hline
    $H $  & $(2, 1/2)$ &  0 &$+$\\
    %$\widetilde{H_{2}}$  & $(2,1/2)$ &  $0$ &$-$\\
    $\eta$ & $(2,1/2)$ & $1-\psi\to 5/2$ &$-$\\
    $S$ & $(1,0)$ & $-2\psi\to 3$ &$+$\\
    $\sigma$ & $(1,0)$ & $\nu+\psi=\eta=1-\psi\to 5/2$&$-$\\
    \hline
    $Q_{L_{i}}$  & $(2,-1/6)$ & $0$& $+$\\
    $\overline{u_{R_{i}}}$ & $(1,-2/3)$ & $0$& $+$\\
    $\overline{d_{R_{i}}}$ & $(1,1/3)$ & $0$& $+$\\
    \hline
    $L_i$  & $(2,-1/2)$ & $-1$& $+$\\
    $\overline{e_{R_i}}$ & $(1,1)$ & $+1$ & $+$ \\
    $\overline{\nu_{R_{i,j}}}$ & $(1,0)$ & $\nu\to 1-2\psi\to 4$&$+$\\
    $\overline{\nu_{R_k}}$ & $(1,0)$ & $\nu_k$&$+$\\
    $\psi_{i}$  & $(1,0)$ & $\psi\to -3/2$ &$-$\\\hline
  \end{tabular}
  \caption{Scalars and left-handed fermions with their respective charges. $Z_2$ to define THDM Type-II~\cite{1303.0168}}
  \label{tab:partcont}
\end{table}
%

>>>>

%Features of the new propossed solution
In general, solutions for light Dirac neutrino masses require a
continuous symmetry to guarantee their Diracness. This symmetry is
usually identified as the $\operatorname{U}(1)_{B-L}$.  Additionally,
ad hoc discrete symmetries are invoked to forbid tree level Dirac or
Majorana mass terms for the light right-handed
neutrinos~\cite{Roncadelli:1983ty,Han:2018zcn,Wang:2017mcy}. 
%tree-level without adhoc symmetries
However, tree-level Dirac type-I seesaw with proper choices for the
$\operatorname{U}(1)_{B-L}$ charges have been shown to be consistent
without require any extra ad-hoc discrete symmetry~\cite{Ma:2014qra}.
%one-loop 
In recent works, it has been shown that even for one-loop Dirac
neutrino masses, it is possible to have $\operatorname{U}(1)_{B-L}$ as
the only extra symmetry beyond the SM~\cite{Calle:2018ovc,Bonilla:2018ynb,Saad:2019bqf}\footnote{
%extra-loops
  For extensions with only extra scalars, minimal solutions have been found with two and three loops~\cite{Saad:2019bqf}}.
%conexion with dark matter
As a bonus in this case, the new scalars and fermions circulating the
loop can be dark matter candidates with the stability of the lightest
of them guaranteed by the very same continuous symmetry.
% remark
We focus here in solutions for the radiative Dirac type-I seesaw with
Majorana mediators, which have only an extra symmetry responsible for the
Diracness of the light neutrinos, the absence of any tree-level
mass, and the existence of a dark sector constituted by the
particles circulating the loop.

%conexion with DM
In fact, another evidence that the SM is not a complete theory is the
missing matter content of the universe, which is known as dark matter
(DM).
% DM as particle
The main proposals that explain DM as a particle are given in
Ref.~\cite{Bertone:2004pz}.
% too many possibilities
Without evidence of DM as a particle there is not a clear path to pin
out the DM properties nor the possible heavier companions of some
extended sector. 
% connect with other phenomena
Linking the dark sector to other specific phenomenology allows to
reduce the arbitrariness in the model building.
% loop
In our construction, the dark sector is related to the heavy sector
responsible of the lightness of the neutrinos and the same symmetry
which guarantees the lightness of the Dirac neutrinos is the
responsible of the stability of the lightest dark particle (LDP).
% new particle prediction
Therefore, the number of specific models is quite restricted.

@@

%
\begin{table}
  \centering
  \begin{tabular}{|l|l|c|c|}
    \hline  
    Particles     & $\left( SU(2)_L, U(1)_Y \right)$ & $U(1)_{L}$& $Z_2$\\ \hline
    $H $  & $(2, 1/2)$ &  0 &$+$\\
    %$\widetilde{H_{2}}$  & $(2,1/2)$ &  $0$ &$-$\\
    $\eta$ & $(2,1/2)$ & $1-\psi\to 5/2$ &$-$\\
    $S$ & $(1,0)$ & $-2\psi\to 3$ &$+$\\
    $\sigma$ & $(1,0)$ & $\nu+\psi=\eta=1-\psi\to 5/2$&$-$\\
    \hline
    $Q_{L_{i}}$  & $(2,-1/6)$ & $0$& $+$\\
    $\overline{u_{R_{i}}}$ & $(1,-2/3)$ & $0$& $+$\\
    $\overline{d_{R_{i}}}$ & $(1,1/3)$ & $0$& $+$\\
    \hline
    $L_i$  & $(2,-1/2)$ & $-1$& $+$\\
    $\overline{e_{R_i}}$ & $(1,1)$ & $+1$ & $+$ \\
    $\overline{\nu_{R_{i,j}}}$ & $(1,0)$ & $\nu\to 1-2\psi\to 4$&$+$\\
    $\overline{\nu_{R_k}}$ & $(1,0)$ & $\nu_k$&$+$\\
    $\psi_{i}$  & $(1,0)$ & $\psi\to -3/2$ &$-$\\\hline
  \end{tabular}
  \caption{Scalars and left-handed fermions with their respective charges. $Z_2$ to define THDM Type-II~\cite{Chiang:2013ixa}}
  \label{tab:partcont}
\end{table}
%
